Comparative safety and effectiveness of Pfizer BA.4-5 versus Sanofi during the spring 2023 COVID-19 booster vaccination programme in England: a matched cohort study in OpenSAFELY-TPP

### 2024-02-27 write-up/manuscript/manuscript.qmd
fix merge errors

IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).

@@

This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes. We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination. For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resultng in reduced power to identify differences between vaccines.

@@

![figure1a](figures_manual/flowchart_cv.png){width="60%"}

@@

```{r figure-contrasts-irr}
#| echo: false
#| warning: false
#| out-height: 50%

@@

plot_irr_table <- contrasts_plot %>%
  ggplot(aes(y=outcome_descr)) +
  geom_text(aes(x=0.5, label=irrCI), hjust=0.5, vjust=0.5, size=9/.pt)+
  #geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0.2, xmax=0.8, fill=patch), alpha=0.2)+
  facet_wrap(vars(cohort_descr), scales="free_y", ncol=1, strip.position = "top")+
  scale_fill_manual(values=c("transparent", "darkgrey"))+
  scale_x_continuous(expand=expansion())+
  #labs(subtitle="\nIncidence\nRate Ratio\n(95% CI)")+
  theme_minimal()+
  theme(
    plot.margin=grid::unit(c(0,0,0,0), "mm"),
    plot.subtitle=element_text(hjust=0.5, vjust=0),
    legend.position="none",
    panel.grid = element_blank(),
    strip.text =element_text(colour="transparent"),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x=element_blank(),
    panel.border = element_blank(),
    panel.spacing = unit(0.5, "lines"),
  )+
  NULL

>>>>

IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).

@@

This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes. We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination. For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resulting in reduced power to identify differences between vaccines.

@@

![figure1a](figures_manual/flowchart_cv.png){width="60%"}
![figure1a](figures_manual/flowchart_cv.png){width="60%"}

@@

```{r figure-contrasts-irr}
#| echo: false
#| warning: false
#| out-height: '50%'

@@

plot_irr_table <- contrasts_plot %>%
  ggplot(aes(y=outcome_descr)) +
  geom_text(aes(x=0.5, label=irrCI), hjust=0.5, vjust=0.5, size=9/.pt)+
  #geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0.2, xmax=0.8, fill=patch), alpha=0.2)+
  facet_wrap(vars(cohort_descr), scales="free_y", ncol=1, strip.position = "top")+
  scale_fill_manual(values=c("transparent", "darkgrey"))+
  scale_x_continuous(expand=expansion())+
  #labs(subtitle="\nIncidence\nRate Ratio\n(95% CI)")+
  theme_minimal()+
  theme(
    plot.margin=grid::unit(c(0,0,0,0), "mm"),
    plot.subtitle=element_text(hjust=0.5, vjust=0),
    legend.position="none",
    panel.grid = element_blank(),
    strip.text =element_text(colour="transparent"),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x=element_blank(),
    panel.border = element_blank(),
    panel.spacing = unit(0.4, "lines"),
  )+
  NULL
